\vspace{-5mm} \fontsize{24pt}{12pt}\selectfont \textbf{What's so Special About Philosophy?}

### 2015-10-25 writeup/draft.tex
network d3 figures complete
parsing algo flags figure

% ------
% Maketitle metadata
\title{\vspace{-5mm}%
	\fontsize{24pt}{12pt}\selectfont
	\textbf{What's So Special About Philosophy?} 
	}	
\author{%
\fontsize{14pt}{14pt}\selectfont
	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
	}
\date{}

@@

Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths\footnote{
nearly $99.6\%$ of first link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
\footnote{
An analysis performed in 2011 by Mat Kelcey posted on his blog (and a smiliar post by Ilmari Karonen) suggest $94.5\%$ of articles end up at Philosophy.
}
lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
\footnote{
cite MIT Tech Review Article: http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/
}
More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
Here we study Wikipedia's First Link Network for insight into how we related topics, ideas, people, objects, and events.

We algorithmically parse every article on the English Wikipedia to construct a map of the First Link Network. 
We then traverse the network, following every possible path to uncover the network's structure.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
basins around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

%========================INTRODUCTION====================================
At no point in history has a larger or more meticulously indexed collection of human knowledge existed.
In amassing such an awe-inspiring collection, we formed an equally impressive web. 
Through the efforts of millions of individuals, working independently,
((cite))
we naturally linked topics, inventions, people, objects, places, and events across space and time.
The web we weaved, and continue to weave, is a wealth of information not only about those notable inventions, 
places, figures, and ideas, but also about \textit{relationships} among them.
Here we study the relationships in this naturally arising web through the hyperlinks connecting one article to another.

We build Wikipedia's First Link Network by following the first link, not in parenthesis, inside the main body of each article in the English version of Wikipedia. 
For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis, 
disregarding pronunciation keys or disambiguations.
We also exclude any links in the side-bar elements, as well as any links to external pages, files, or WikiMedia 
projects outside of Wikipedia (such as Wiktionary).
This procedure corresponds to the original claim in 2008 about the percentage of pages with first links to 
"Philosophy"\footnote{
https://en.wikipedia.org/wiki/Wikipedia:GettingtoPhilosophy
}
\footnote{
http://xkcd.com/903/
}
\footnote{
https://www.reddit.com/r/pics/comments/gpdhb/trythiswikipediamindfk/
}

@@

To map Wikipedia's First Link Network, we use the freely-available XML dump of the English Wikipedia. 
Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
eliminating any statistical error due to sampling.
We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
and disambiguations.
Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
article and another.

@@

While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article.\footnote{
Approaches using existing libraries led to several bugs 
including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
as well as compatibility with other libraries used to parse the XML.}
Our parsing algorithm aimed to: 
1. squash the initial bugs 
2. eliminate the need for several passes. 
To process an article with one pass, we developed a hierarchical system of flags:
\begin{itemize}
    \item[Flag 1:] inside Wikimedia template?\\
        * \bt{trigger}: \{\{
    \begin{itemize}
        \item[Flag 2:] inside <ref>, <div>?
        \begin{itemize}
            \item[Flag 3:] inside ()?
        \end{itemize}
    \end{itemize}
\end{itemize}

\red{[[better graphic to come]]}

@@

@@

\begin{figure}[H]
\centering
\caption{highest ranking 3-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
    \end{subfigure}
\end{figure}
((diagram of cycles beyond length 10))

@@

@@

\begin{figure}[H]
\centering
    \caption{Longest Path Lengths}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_1k_path_length.png}
    \end{subfigure}
\end{figure}

@@

\begin{figure}[H]
\centering
    \caption{Top Funnels}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
    \end{subfigure}
\end{figure}

@@

%========================Style Guide====================================

>>>>

% ------
% Maketitle metadata
\title{\vspace{-5mm}%
	\fontsize{24pt}{12pt}\selectfont
	\textbf{What's so Special About Philosophy?} 
	}	
\author{%
\fontsize{14pt}{14pt}\selectfont
	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
	}
\date{}

@@

Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths\footnote{
nearly $99.6\%$ of first link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
\footnote{
An analysis performed in 2011 by Mat Kelcey posted on his blog (and a smiliar post by Ilmari Karonen) suggest $94.5\%$ of articles end up at Philosophy.
}
lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
\footnote{
cite MIT Tech Review Article: http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/
}
More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.

We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
In a novel approach to uncover structure, we traverse every possible path through the network, 
measuring the accumulation of first links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
basins around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

%========================INTRODUCTION====================================
At no point in history has a larger or more meticulously indexed collection of human knowledge existed.
In amassing such an awe-inspiring collection, we formed an equally impressive web. 
Through the efforts of millions of individuals, working independently,
((cite))
we naturally linked topics, inventions, people, objects, places, and events across space and time.
The web we weaved, and continue to weave, is a wealth of information not only about those notable inventions, 
places, figures, and ideas, but also about \textit{relationships} among them.
We study the relationships in this naturally arising web through the hyperlinks connecting one article to another.

We build Wikipedia's First Link Network by following the first link, not in parenthesis, inside the main body of each article in the English version of Wikipedia. 
For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis, 
disregarding pronunciation keys or disambiguations.
We also exclude any links in the side-bar elements, as well as any links to external pages, files, or WikiMedia 
projects outside of Wikipedia (such as Wiktionary).
This procedure corresponds to the original claim in 2008 about the percentage of pages with first links to 
"Philosophy."\footnote{
https://en.wikipedia.org/wiki/Wikipedia:GettingtoPhilosophy
}
\footnote{
http://xkcd.com/903/
}
\footnote{
https://www.reddit.com/r/pics/comments/gpdhb/trythiswikipediamindfk/
}

@@

To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
eliminating any statistical error due to sampling.
We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
and disambiguations.
Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
article and another.

@@

While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article.\footnote{
Approaches using existing libraries led to several bugs 
including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
as well as compatibility with other libraries used to parse the XML.}
Our parsing algorithm aimed to: 
1) accurately identify the first link among other page elements
2) efficiently do so, without the need for several passes through the data.
To process an article with one pass, we developed a hierarchical system of flags:

\begin{figure}[H]
\centering
\caption{parsing algorithm for XML}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/flags.pdf}    
    \end{subfigure}
\end{figure}

@@

\begin{figure}[H]
\centering
\caption{highest ranking 2-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
    \end{subfigure}
\end{figure}

@@

\begin{figure}[H]
\centering
\caption{highest ranking 3-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
    \end{subfigure}
\end{figure}
((diagram of cycles beyond length 10))

@@

\begin{figure}[H]
\centering
\caption{highest ranking 3-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/basins.pdf}
    \end{subfigure}
\end{figure}

@@

\begin{figure}[H]
\centering
    \caption{Path Length Distribution}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/path_lengths_boxplot.png}
    \end{subfigure}
\end{figure}
\begin{figure}[H]
\centering
    \caption{Longest Path Lengths}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_1k_path_length.png}
    \end{subfigure}
\end{figure}

@@

\begin{figure}[H]
\centering
    \caption{Top Funnels}
    \begin{subfigure}[b]{0.9\textwidth}
        \includegraphics[width=\textwidth]{graphics/funnels.pdf}
    \end{subfigure}
\end{figure}

@@

\section{Supplementary}

\begin{figure}[H]
\centering
\caption{highest ranking 2-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
    \end{subfigure}
\end{figure}
\begin{figure}[H]
\centering
    \caption{Top Funnels}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
    \end{subfigure}
\end{figure}
%========================Style Guide====================================
